Model Finding for Recursive Functions in SMT

### 2016-01-26 ijcar2016/body.tex
tuned

\cvc implements a few optimizations designed to help finding finite models of
$\conv(\varphi)$.
Like other systems,
the finite model finding capability of \cvc incrementally fixes bounds on the cardinalities of uninterpreted types
and increases these bounds until it encounters a model.
When multiple types are present, it uses a fairness scheme that
bounds the sum of cardinalities of all uninterpreted types~\cite{reynolds2013finite}.
For example, if a signature has two uninterpreted types $\tau_1$ and $\tau_2$,
it will first search for models where 
$\left\vert\tau_1\right\vert + \left\vert\tau_2\right\vert$ is at most $2$,
then $3$, $4$, and so on.
To accelerate the search for models,
we implemented an optimization based on statically inferring \emph{monotonic} types.
A monotonic type is one in which models can always be extended with additional
elements of that type \cite{blanchette-krauss-2011,claessen-et-al-2011}.
%%% I find this part more confusing than helping, since we use monotonicity to merge
%%% sorts (or at least their cardinality bounds), whereas Paradox and iProver use
%%% it to separate sorts! Our use of monotonicity is more related to Nitpick's. --jb
%%%
% korovin-2013
%Inference of monotonic types has been show in previous work~
%to be an important component of finite model finding systems.
Types $\fargtype{\con{f}}$ introduced by our translation $\conv$ are monotonic,
because $\teq$ is never used directly on such types \cite{claessen-et-al-2011}.
%
%The finite model finding capability of
\cvc takes advantage of this by fixing the bounds for all 
monotonic types simultaneously.
That is, if $\tau_1$ and $\tau_2$ are inferred monotonic (whether they are
introduced by our translation or are present in the original problem),
the solver fixes the bound for \relax{both} types to be $1$, then $2$, and so on.
This scheme allows the solver greater flexibility compared with the default scheme,
and comes with no loss of generality with respect to models,
since monotonic types can always be extended to have equal cardinalities.

By default, \cvc uses techniques to minimize the number of literals it
considers when constructing propositional satisfying assignments for
formulas~\cite{relevancy2007}. However, we have found such techniques degrade
performance for finite model finding on problems having recursive functions
that are defined by cases. For this reason, we prefer disabling such
techniques for problems coming from our translation.

>>>>

\cvc implements a few optimizations designed to help finding finite models of
$\conv(\varphi)$.
Like other systems,
the finite model finding capability of \cvc incrementally fixes bounds on the cardinalities of uninterpreted types
and increases these bounds until it encounters a model.
When multiple types are present, it uses a fairness scheme that
bounds the sum of cardinalities of all uninterpreted types~\cite{reynolds2013finite}.
For example, if a signature has two uninterpreted types $\tau_1$ and $\tau_2$,
it will first search for models where 
$\left\vert\tau_1\right\vert + \left\vert\tau_2\right\vert$ is at most $2$,
then $3$, $4$, and so on.
To accelerate the search for models,
we implemented an optimization based on statically inferring \emph{monotonic} types.
A monotonic type is one in which models can always be extended with additional
elements of that type \cite{blanchette-krauss-2011,claessen-et-al-2011}.
%%% I find this part more confusing than helping, since we use monotonicity to merge
%%% sorts (or at least their cardinality bounds), whereas Paradox and iProver use
%%% it to separate sorts! Our use of monotonicity is more related to Nitpick's. --jb
%%%
% korovin-2013
%Inference of monotonic types has been show in previous work~
%to be an important component of finite model finding systems.
Types $\fargtype{\con{f}}$ introduced by our translation $\conv$ are monotonic,
because $\teq$ is never used directly on such types \cite{claessen-et-al-2011}.
%
%The finite model finding capability of
\cvc takes advantage of this by fixing the bounds for all 
monotonic types simultaneously.
That is, if $\tau_1$ and $\tau_2$ are inferred monotonic (whether they are
present in the original problem or introduced by our translation),
the solver fixes the bound for \relax{both} types to be $1$, then $2$, and so on.
This scheme allows the solver greater flexibility compared with the default scheme,
and comes with no loss of generality with respect to models,
since monotonic types can always be extended to have equal cardinalities.

By default, \cvc uses techniques to minimize the number of literals it
considers when constructing propositional satisfying assignments for
formulas~\cite{relevancy2007}. However, we have found such techniques degrade
performance for finite model finding on problems having recursive functions
that are defined by cases. For this reason, we disable the techniques for
problems produced from our translation.
